\papertitle

### 2014-03-20 final-report/paper.tex
intro

\item A view for cluster utilization and operator performance for each query fragment. Each fragment is composed of
multiple sequential database operators running on the same worker. Upon clicking on a collapsed fragment node in the
query plan DAG, the node expands to display the operators comprising the fragment and renders the fragment execution view.
The fragment execution view visualizes the utilization of the cluster by showing whether each worker is busy working
on the selected fragment. It also breaks down the work done by each worker into operators composing the fragment.

\item A view for communication between workers. A directed link between two fragments in the physical query plan DAG
represents source workers sending tuples to destination workers. Upon clicking on a link, the visualization renders the
worker communication view. This shows the aggregate number of tuples exchanged between workers as well as a time series
charting the number of tuples exchanged over time.

@@

The \graph view as pictured in Figure~\ref{fig:graph} allows the user to selectively browse different performance visualizations tied to a given query execution. The \graph view provides as the name indicates, a visualization of the physical query plan. A query plan is an ordered set of steps used to access data in database management systems. The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next. Each query fragment is a collection of query operators. One can think of a query fragment as a job unit that gets scheduled on a cluster of machines. Consequently transitioning from one query fragment to another generally leads to all-to-all communication between worker nodes on the cluster.

@@

    \item \textbf{Empty selection:} by default if no fragments or edges are selected in the \graph view, an \overall visualization is rendered in the performance panel. This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets rendered in the performance panel. The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules. When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker task schedule in the performance window with the corresponding query operator in the browser window.
    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-worker communication from one query fragment to the next.

@@

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/utilization_chart}
  \caption{The utilization chart. The user can brush on either the small graph on top or the bottom graph. Brushing on the second chart zooms the chart.}
  \label{fig:utilization_chart}
\end{figure}

@@

\subsubsection{Worker communication view}

The Worker Communication view consists of two charts: the aggregate communication
chart and the time series chart. The aggregate communication chart is a matrix (Figure~\ref{fig:network}) representing
a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
by worker $m$ to worker $n$ during the query execution. We chose to use the colors from color brewer\cite{harrower2003colorbrewer} and interpolated the colors between the steps using the \emph{CIE L*a*b*} interpolation and lightness correction \cite{chromajsaisch}. The colors work in grayscale and can be interpreted by color blind users.

The matrix chart representing aggregate communication between workers help uncover data skew
issues, as discussed in Section~\ref{sec:eval}.

@@

\begin{figure}[!ht]
  \includegraphics[width=\columnwidth]{images/networkVis2}
  \caption{Worker communication time series chart. The user can select multiple cells in the matrix to compare the
  corresponding time series.}
  \label{fig:ts}
\end{figure}

>>>>

\item \emph{How does the cluster visualization vary for each fragment and operator?} 
%A view for cluster utilization and operator performance for each query fragment. 
In \emph{Myria}, a fragment is composed of multiple sequential database operators running on the same worker. 
A programmer will want to understand the cluster utilization at a given time as well as how much time is being spent
executing an operator. \system provides a \fragment view to aid this understanding. 
%Upon clicking on a collapsed fragment node in the
%query plan DAG, the node expands to display the operators comprising the fragment and renders the fragment execution view.
The \fragment view visualizes the utilization of the cluster by showing whether each worker is busy working
on the selected fragment. It also breaks down the work done by each worker into operators composing the fragment.

\item A view for communication between workers. A directed link between two fragments in the physical query plan DAG
represents source workers sending tuples to destination workers. Upon clicking on a link, the visualization renders the
\network view. This shows the aggregate number of tuples exchanged between workers as well as a time series
charting the number of tuples exchanged over time.

@@

The \graph view as pictured in Figure~\ref{fig:graph} allows the user to selectively browse different performance visualizations tied to a given query execution. The \graph view provides as the name indicates, a visualization of the physical query plan. A query plan is an ordered set of steps used to access data in database management systems. The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-\network from one query fragment to the next. Each query fragment is a collection of query operators. One can think of a query fragment as a job unit that gets scheduled on a cluster of machines. Consequently transitioning from one query fragment to another generally leads to all-to-all communication between worker nodes on the cluster.

@@

    \item \textbf{Empty selection:} by default if no fragments or edges are selected in the \graph view, an \overall visualization is rendered in the performance panel. This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets rendered in the performance panel. The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules. When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker task schedule in the performance window with the corresponding query operator in the browser window.
    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-\network from one query fragment to the next.

@@

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/utilization_chart}
  \caption{The utilization chart.}
  \label{fig:utilization_chart}
\end{figure}

@@

\subsubsection{\network view}

The \network view consists of two charts: the aggregate communication
chart and the time series chart. The aggregate communication chart is a matrix (Figure~\ref{fig:network}) representing
a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
by worker $m$ to worker $n$ during the query execution. We chose to use the colors from color brewer\cite{harrower2003colorbrewer} and interpolated the colors between the steps using the \emph{CIE L*a*b*} interpolation and lightness correction \cite{chromajsaisch}. The colors work in grayscale and can be interpreted by color blind users.

The matrix chart representing aggregate communication between workers help uncover data skew
issues, as discussed in \S \ref{sec:eval}.

% We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
%for the aggregate communication chart. But feedback during the progress presentation session convinced us that
%a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
%part of future work.

\todo{is this figure necessary or is the figure in the evaluation section enough?}

\begin{figure}[!ht]
  \includegraphics[width=\columnwidth]{images/networkVis1}
  \caption{\network matrix chart. Each cell represents the number of exchanged tuples between a
  pair of workers.}
  \label{fig:matrix}
\end{figure}

@@

\begin{figure}[!ht]
  \includegraphics[width=\columnwidth]{images/networkVis2}
  \caption{\network time series chart. The user can select multiple cells in the matrix to compare the
  corresponding time series.}
  \label{fig:ts}
\end{figure}
